---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpmr

<!-- badges: start -->
[![R-CMD-check](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/psychelzh/cpmr/graph/badge.svg)](https://app.codecov.io/gh/psychelzh/cpmr)
[![CRAN status](https://www.r-pkg.org/badges/version/cpmr)](https://CRAN.R-project.org/package=cpmr)
<!-- badges: end -->

The cpmr package is designed for connectome predictive modeling (CPM) in R.
Its primary workflow is a native, matrix-first API that keeps CPM-specific
training, resampling, and leakage-safe preprocessing inside package-controlled
code. This package relies on [Rfast](https://CRAN.R-project.org/package=Rfast)
for row-oriented calculation.

## Installation

You can install the released version of cpmr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cpmr")
```

Or you can install the development version of cpmr from [r-universe](https://psychelzh.r-universe.dev) with:

``` r
install.packages("cpmr", repos = c("https://psychelzh.r-universe.dev", getOption("repos")))
```

## Native workflow

Shape your connectivity matrix as a subjects-by-edges matrix, where each row
contains the edge vector for one subject, and pair it with a behavioral vector.
The most direct entry point is `cpm_fit()`.

```{r example}
library(cpmr)

withr::local_seed(123)
conmat <- matrix(rnorm(100 * 1000), nrow = 100)
behav <- rnorm(100)
fit_obj <- cpm_fit(
  conmat = conmat,
  behav = behav
)

fit_obj
summary(fit_obj)
```

Cross-validated resampling uses the matching native helper
`cpm_fit_resamples()`:

```{r resample}
resample_obj <- cpm_fit_resamples(
  conmat = conmat,
  behav = behav,
  kfolds = 5
)

collect_metrics(resample_obj)
collect_edges(resample_obj, format = "index")
```

## Choosing a path

`cpmr` now treats the native workflow as the primary package story:

- use `cpm_fit()` and `cpm_fit_resamples()` for most real CPM analyses;
- use `fit(cpm_spec(), ...)` and `fit_resamples(cpm_spec(), ...)` when you want
  the lower-level specification object directly;
- treat external workflow adapters as optional future interoperability layers,
  not as the default path for heavy resampling or tuning.

Why this matters: CPM often needs leakage-safe fold-local preprocessing and can
benefit from future fold-level caching or threshold-specific optimization.
Those workloads fit native `cpmr` runners better than a generic orchestration
layer.

## Code of Conduct

Please note that the cpmr project is released with a [Contributor Code of Conduct](https://psychelzh.github.io/cpmr/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
