---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpmr

<!-- badges: start -->
[![R-CMD-check](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/psychelzh/cpmr/graph/badge.svg)](https://app.codecov.io/gh/psychelzh/cpmr)
[![CRAN status](https://www.r-pkg.org/badges/version/cpmr)](https://CRAN.R-project.org/package=cpmr)
<!-- badges: end -->

The cpmr package implements connectome predictive modeling (CPM) in R.
Its primary workflow is native and matrix-first: you pass a subject-by-edge
matrix together with a behavioral vector, and `cpmr` handles CPM-specific
training, resampling, and leakage-safe preprocessing inside package code.
This package relies on [Rfast](https://CRAN.R-project.org/package=Rfast) for
row-oriented calculation.

## Installation

You can install the released version of cpmr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cpmr")
```

Or you can install the development version of cpmr from [r-universe](https://psychelzh.r-universe.dev) with:

``` r
install.packages("cpmr", repos = c("https://psychelzh.r-universe.dev", getOption("repos")))
```

## Native workflow

Shape your connectivity matrix as a subjects-by-edges matrix, where each row
contains the edge vector for one subject, and pair it with a behavioral vector.
The native workflow uses `spec()` together with `cpm()`.

```{r example}
library(cpmr)

withr::local_seed(123)
conmat <- matrix(rnorm(100 * 1000), nrow = 100)
behav <- rnorm(100)
s <- spec()
resample_obj <- cpm(conmat = conmat, behav = behav, spec = s, resamples = 5)

summary(resample_obj)
head(resample_obj$predictions)
```

`spec()` can also express richer CPM variants without leaving the native
API surface:

```{r rich-spec}
rich_spec <- spec(
  selection = cpm_selection_cor(
    method = "spearman",
    criterion = "absolute",
    level = 0.1
  ),
  construction = cpm_construction_summary(
    sign_mode = "net",
    weight_scale = 0.03
  ),
  model = cpm_model_lm()
)

cpm(conmat = conmat, behav = behav, spec = rich_spec, resamples = 5)$predictions |>
  head()
```

This keeps the main CPM stages visible while grouping naturally paired options
into small helpers:

- `selection = cpm_selection_cor(...)` defines how edges are screened;
- `construction = cpm_construction_summary(...)` defines how screened edges
  become CPM-derived predictors;
- `model = cpm_model_lm()` defines the outcome model fitted on those
  predictors.

By default, `spec()` keeps edge standardization turned off so the native
workflow stays close to the classic CPM path of screening edges, constructing
summary features, and fitting the outcome model. If you want fold-local edge
z-scoring, opt in with
`cpm_construction_summary(standardize_edges = TRUE)`.

If you are new to CPM, the key idea behind the current summary-based
construction is:

- `positive` edges are edges whose screening association with the outcome is
  positive and passes the threshold;
- `negative` edges are edges whose screening association with the outcome is
  negative and passes the threshold;
- CPM then collapses those screened edge sets into subject-level summary
  features before fitting the outcome model.

With `sign_mode = "separate"`, `cpmr` keeps the positive and negative summaries
as separate predictors and reports three prediction streams:

- `joint`, which fits one model with both summaries together;
- `positive`, a positive-only diagnostic stream;
- `negative`, a negative-only diagnostic stream.

With `sign_mode = "net"`, `cpmr` constructs a single
`net_summary = positive_summary - negative_summary` feature and returns one
`net` prediction stream.

`cpm()` is resample-first: it runs CPM under the requested assessment plan and
returns out-of-fold predictions together with the fold structure.

```{r resample}
summary(resample_obj)
head(resample_obj$predictions)
dim(resample_obj$edges)
head(resample_metrics(resample_obj))
```

`summary(resample_obj)` gives the default aggregate report, with pooled
out-of-fold error metrics shown first and correlations reported as
supplementary statistics. Use `resample_metrics(resample_obj)` when you want
pooled or fold-wise metric tables directly.

For `cpm()`, `resamples = NULL` means leave-one-out resampling,
`resamples = 5` means 5-fold resampling, and a list passed to `resamples`
enforces a manual assessment-fold partition.

## Choosing a path

`cpmr` treats this native workflow as the primary package path:

- use `cpm(..., spec = spec())` for native CPM analyses;
- keep the `spec()` object around when you want an explicit, reusable
  parameter object.

Why this matters: CPM often needs leakage-safe fold-local preprocessing and can
benefit from future fold-level caching or threshold-aware optimization. Those
workloads fit native `cpmr` runners better than a generic orchestration layer.

## Code of Conduct

Please note that the cpmr project is released with a [Contributor Code of Conduct](https://psychelzh.github.io/cpmr/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
