---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpmr

<!-- badges: start -->
[![R-CMD-check](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/psychelzh/cpmr/graph/badge.svg)](https://app.codecov.io/gh/psychelzh/cpmr)
[![CRAN status](https://www.r-pkg.org/badges/version/cpmr)](https://CRAN.R-project.org/package=cpmr)
<!-- badges: end -->

`cpmr` brings connectome-based predictive modeling (CPM) into the
tidymodels ecosystem. It provides:

- a `parsnip` model specification through `cpm_reg()`;
- workflow-friendly fitting with formulas or recipes;
- tuning support for CPM thresholds;
- CPM-aware metrics such as `cpm_cor()` and `cpm_spearman()`;
- edge inspection helpers through `collect_edges()`.

The expected predictor layout is one subject per row and one edge per
column.

## Installation

You can install the released version of cpmr from
[CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cpmr")
```

Or you can install the development version from GitHub with:

``` r
pak::pak("psychelzh/cpmr")
```

## Fit A CPM Model

```{r fit-example}
library(cpmr)
library(parsnip)

withr::local_seed(123)

n <- 80
p <- 40
x <- matrix(rnorm(n * p), nrow = n, ncol = p)
y <- x[, 1] - x[, 2] + 0.5 * x[, 3] + rnorm(n, sd = 0.5)

dat <- as.data.frame(x)
names(dat) <- paste0("edge_", seq_len(p))
dat$y <- y

spec <- cpm_reg(
  thresh_method = "sparsity",
  thresh_level = 0.2,
  network = "both"
)

fit_obj <- fit(spec, y ~ ., data = dat[1:60, ])
predict(fit_obj, dat[61:80, ])
collect_edges(fit_obj)[1:12, , drop = FALSE]
```

## Use With Workflows

```{r workflow-example}
suppressPackageStartupMessages({
  library(recipes)
  library(workflows)
})

wf <- workflow() |>
  add_recipe(
    recipe(y ~ ., data = dat[1:60, ]) |>
      step_normalize(all_predictors())
  ) |>
  add_model(spec)

wf_fit <- fit(wf, data = dat[1:60, ])
predict(wf_fit, dat[61:80, ])
```

## Tune CPM Thresholds

```{r tune-example}
library(rsample)
library(tune)
library(yardstick)

folds <- vfold_cv(dat[1:60, ], v = 3)

tuned_spec <- cpm_reg(
  thresh_method = "sparsity",
  thresh_level = tune(),
  network = "both"
)

tune_res <- workflow() |>
  add_formula(y ~ .) |>
  add_model(tuned_spec) |>
  tune_grid(
    resamples = folds,
    grid = tibble::tibble(thresh_level = c(0.15, 0.25)),
    metrics = metric_set(rmse, cpm_cor)
  )

collect_metrics(tune_res)
```

## Design Notes

CPM is a protocol, not just a single regression call. In `cpmr`, edge
selection, network aggregation, and model fitting are all kept inside
each resampling split so that feature selection stays independent from
assessment data.

## Code Of Conduct

Please note that the cpmr project is released with a [Contributor Code
of Conduct](https://psychelzh.github.io/cpmr/CODE_OF_CONDUCT.html). By
contributing to this project, you agree to abide by its terms.
