---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpmr

<!-- badges: start -->
[![R-CMD-check](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/psychelzh/cpmr/graph/badge.svg)](https://app.codecov.io/gh/psychelzh/cpmr)
[![CRAN status](https://www.r-pkg.org/badges/version/cpmr)](https://CRAN.R-project.org/package=cpmr)
<!-- badges: end -->

The cpmr package implements connectome predictive modeling (CPM) in R.
Its primary workflow is native and matrix-first: you pass a subject-by-edge
matrix together with a behavioral vector, and `cpmr` handles CPM-specific
training, resampling, and leakage-safe preprocessing inside package code.
This package relies on [Rfast](https://CRAN.R-project.org/package=Rfast) for
row-oriented calculation.

## Installation

You can install the released version of cpmr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cpmr")
```

Or you can install the development version of cpmr from [r-universe](https://psychelzh.r-universe.dev) with:

``` r
install.packages("cpmr", repos = c("https://psychelzh.r-universe.dev", getOption("repos")))
```

## Native workflow

Shape your connectivity matrix as a subjects-by-edges matrix, where each row
contains the edge vector for one subject, and pair it with a behavioral vector.
The native workflow uses `cpm_spec()` together with `fit()`.

```{r example}
library(cpmr)

withr::local_seed(123)
conmat <- matrix(rnorm(100 * 1000), nrow = 100)
behav <- rnorm(100)
fit_obj <- fit(cpm_spec(), conmat = conmat, behav = behav)

fit_obj
summary(fit_obj)
```

`cpm_spec()` can also express richer CPM variants without leaving the native
API surface:

```{r rich-spec}
rich_spec <- cpm_spec(
  screen = cpm_screen(
    association = "spearman",
    threshold = cpm_threshold("effect_size", level = 0.1)
  ),
  feature_space = "net",
  weighting = cpm_weighting("sigmoid", scale = 0.03),
  model = cpm_model_lm()
)

fit(rich_spec, conmat = conmat, behav = behav)$predictions |>
  head()
```

This keeps the main CPM choices visible while grouping naturally paired options
like screening, edge weighting, and the outcome model into small helper
objects.

Cross-validated resampling uses `fit_resamples()` with the same specification
object:

```{r resample}
resample_obj <- fit_resamples(cpm_spec(), conmat = conmat, behav = behav, kfolds = 5)

summary(resample_obj)
head(resample_obj$predictions)
dim(resample_obj$edges)
head(resample_metrics(resample_obj))
```

`summary(resample_obj)` gives the default aggregate report, with pooled
out-of-fold error metrics shown first and correlations reported as
supplementary statistics. Use `resample_metrics(resample_obj)` when you want
pooled or fold-wise metric tables directly.

## Choosing a path

`cpmr` treats this native workflow as the primary package path:

- use `fit(cpm_spec(), ...)` and `fit_resamples(cpm_spec(), ...)` for native
  CPM analyses;
- keep the `cpm_spec()` object around when you want an explicit, reusable
  parameter object.

Why this matters: CPM often needs leakage-safe fold-local preprocessing and can
benefit from future fold-level caching or threshold-aware optimization. Those
workloads fit native `cpmr` runners better than a generic orchestration layer.

## Code of Conduct

Please note that the cpmr project is released with a [Contributor Code of Conduct](https://psychelzh.github.io/cpmr/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
