---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpmr

<!-- badges: start -->
[![R-CMD-check](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/psychelzh/cpmr/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/psychelzh/cpmr/branch/main/graph/badge.svg)](https://app.codecov.io/gh/psychelzh/cpmr?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/cpmr)](https://CRAN.R-project.org/package=cpmr)
<!-- badges: end -->

The cpmr package is specifically designed for the analysis of the connectome predictive modeling (CPM) method in R. This package relies on [Rfast](https://CRAN.R-project.org/package=Rfast) to do row oriented calculation.

## Installation

You can install the released version of cpmr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cpmr")
```

Or you can install the development version of cpmr from [r-universe](https://psychelzh.r-universe.dev) with:

``` r
install.packages("cpmr", repos = c("https://psychelzh.r-universe.dev", getOption("repos")))
```

## Example

It is very simple to use this package. Just shape your connectivity matrix as a subjects by edges matrix, i.e., each row contains the correlation matrix (removed diagonal and duplicated values, e.g., lower triangular data) for each subject, and your behavior data a vector and feed them in `cpm()` function.

```{r example}
library(cpmr)

withr::local_seed(123)
conmat <- matrix(rnorm(100 * 1000), nrow = 100)
behav <- rnorm(100)
res <- cpm(conmat, behav, kfolds = 10)
plot(res$real, res$pred[, "both"], xlab = "Real", ylab = "Predicted")
```
