---
title: "Getting Started with cpmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with cpmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is CPM?

Connectome Predictive Modeling (CPM) is a method for predicting individual
differences in behavior from brain functional connectivity data. The core idea
is straightforward:

1. Measure how strongly each **brain edge** (pair of regions) correlates with
   the behavior of interest — across subjects.
2. Keep only edges that pass a significance threshold (the "predictive network").
3. Summarize those edges into a single score per subject (e.g., by summing their
   connectivity values).
4. Use that score to predict behavior via linear regression.

`cpmr` implements this pipeline in a tidy, cross-validation–friendly way.
This vignette walks you through a complete workflow from raw data to interpreted
results.

## Setup

```{r}
library(cpmr)
```

## Prepare Your Data

CPM takes three inputs:

| Argument | What it is | Shape |
|---|---|---|
| `conmat` | Connectivity matrix — one row per subject, one column per edge | `n × p` matrix |
| `behav` | Behavioral score — one value per subject | length-`n` vector |
| `covariates` | Nuisance variables to regress out (optional) | `n × q` matrix |

Below we simulate a small dataset for illustration. In real analyses, `conmat`
would come from your neuroimaging pipeline (e.g., a Fisher-z transformed
correlation matrix for each subject), and `behav` from a cognitive or clinical
measure.

```{r}
withr::local_seed(123)
n <- 80   # subjects
p <- 200  # brain edges (in practice, tens of thousands)

# Simulate connectivity data
conmat <- matrix(rnorm(n * p), nrow = n, ncol = p)

# Make behavior correlated with the first 12 edges so CPM has signal to detect
behav <- rowMeans(conmat[, 1:12, drop = FALSE]) + rnorm(n, sd = 0.5)

# One nuisance covariate (e.g., head motion, age)
covariates <- matrix(rnorm(n), ncol = 1)
```

## Configure the Model

Before fitting, create a model specification with `cpm_spec()`. This bundles
all algorithm choices into a single reusable object.

```{r}
spec <- cpm_spec(
  thresh_method = "alpha",  # select edges by p-value
  thresh_level  = 0.05,     # keep edges with p < 0.05
  bias_correct  = TRUE      # center and scale connectivity features using training-set statistics
)
```

`thresh_method = "alpha"` is the classic CPM approach: include any edge whose
correlation with behavior is significant at the chosen alpha level. You can
also use `thresh_method = "sparsity"` to keep the top-*k* % of edges by
correlation magnitude (controlling network sparsity rather than significance).

## Step 1 — Fit on the Full Sample

`fit()` trains a CPM model using all subjects at once. This is useful for
**inspecting the selected network** or for hypothesis-driven analyses where
cross-validation is not needed.

```{r}
fit_obj <- fit(
  spec,
  conmat     = conmat,
  behav      = behav,
  covariates = covariates,
  return_edges = "sum",   # store edge selection counts (edges × networks matrix)
  na_action  = "exclude"  # silently drop subjects with any missing value
)

fit_obj
summary(fit_obj)
```

`summary()` reports the correlation between predicted scores and observed behavior.
It returns a single row with three columns — `both`, `pos`, and `neg` — corresponding
to the combined network, the positive network, and the negative network respectively.
The default method is Pearson; pass `method = "spearman"` to use Spearman instead.

## Step 2 — Cross-Validated Resampling

A single in-sample fit is **optimistic**: the model is evaluated on the same
data it was trained on, so the performance estimate is inflated. To get an
honest generalization estimate, use **k-fold cross-validation** via
`fit_resamples()`.

In k-fold CV, the data are split into *k* non-overlapping folds. Each fold
takes a turn as the held-out test set while the model trains on the remaining
*k − 1* folds. This way, every subject contributes exactly one out-of-sample
prediction.

```{r}
set.seed(123)
res_obj <- fit_resamples(
  spec,
  conmat     = conmat,
  behav      = behav,
  covariates = covariates,
  kfolds     = 5,           # 5-fold cross-validation
  return_edges = "sum",
  na_action  = "exclude"
)

res_obj
```

## Step 3 — Extract Results

`cpmr` provides three accessor functions to pull results out of a resampling
object.

### Performance metrics

`collect_metrics()` returns the predictive performance for each fold and
network, expressed as the correlation between predicted and observed behavior.

```{r}
metrics <- collect_metrics(res_obj)
head(metrics)
```

Each row is one fold. The columns `both`, `pos`, and `neg` give the
out-of-sample Pearson *r* for the combined, positive, and negative networks
respectively. `n_assess` records how many subjects were in that fold's
held-out set. Averaging across folds gives you the overall cross-validated
performance of the model.

### Out-of-sample predictions

`collect_predictions()` returns the raw predicted value for every subject in
every fold they were held out.

```{r}
predictions <- collect_predictions(res_obj)
head(predictions)
```

You can use these to compute custom metrics, plot predicted vs. observed, or
inspect individual subjects.

### Selected edges

`collect_edges()` returns a matrix of edge selection counts across folds
(for `return_edges = "sum"`): each cell records how many folds selected that
edge. Higher counts indicate more consistently selected edges.

```{r}
edges <- collect_edges(res_obj)
str(edges)
```

Consistently selected edges across folds are the most reliable candidates for
a "predictive network".

## Advanced: Custom Resamples

By default, `fit_resamples()` generates random k-fold splits internally. You
can override this with your own partition scheme — useful when you need
**site-stratified**, **family-aware**, or **repeated** splits.

Pass a list where each element is a vector of **held-out row indices** for that
fold. `cpmr` validates that the sets (a) are non-overlapping and (b) together
cover every complete-case subject exactly once. When `na_action = "exclude"`,
indices refer to rows in the complete-case data, not the original data.

```{r}
# Manual 5-fold split: subjects 1–16, 17–32, …
custom_resamples <- list(
  1:16,
  17:32,
  33:48,
  49:64,
  65:80
)

custom_res_obj <- fit_resamples(
  spec,
  conmat    = conmat,
  behav     = behav,
  resamples = custom_resamples,
  return_edges = "none"   # skip edge storage to save memory
)

collect_metrics(custom_res_obj)
```

## Next Steps

- **Covariate leakage**: If your covariates include site, age, or motion, read
  the leakage-focused vignette to see how `cpmr` prevents data leakage during
  covariate regression.
- **Reporting**: `tidy()` converts a `cpm` object into a tidy data frame
  suitable for tables and plots; `summary()` gives a quick console summary.
