---
title: "Getting Started with cpmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with cpmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr` implements connectome predictive modeling (CPM) for behavioral prediction
from functional connectivity features.

This vignette gives a minimal workflow for:

- preparing inputs;
- fitting a single CPM model;
- running cross-validated resampling;
- collecting and inspecting outputs.

## Data Shape

CPM in `cpmr` expects:

- `conmat`: a numeric matrix with rows as subjects and columns as edges;
- `behav`: a numeric vector with one value per subject;
- optional `covariates`: numeric matrix/vector with one row per subject.

```{r}
library(cpmr)

withr::local_seed(123)
n <- 80
p <- 200
conmat <- matrix(rnorm(n * p), nrow = n, ncol = p)
behav <- rowMeans(conmat[, 1:12, drop = FALSE]) + rnorm(n, sd = 0.5)
covariates <- matrix(rnorm(n), ncol = 1)
```

## Single-Fit Workflow

Use `fit()` when you want one in-sample fitted object.

```{r}
spec <- cpm_spec(
  thresh_method = "alpha",
  thresh_level = 0.05,
  bias_correct = TRUE
)

fit_obj <- fit(
  spec,
  conmat = conmat,
  behav = behav,
  covariates = covariates,
  return_edges = "sum",
  na_action = "exclude"
)

fit_obj
summary(fit_obj)
```

## Resampling Workflow

Use `fit_resamples()` for cross-validated or custom resampling workflows.

```{r}
set.seed(123)
res_obj <- fit_resamples(
  spec,
  conmat = conmat,
  behav = behav,
  covariates = covariates,
  kfolds = 5,
  return_edges = "sum",
  na_action = "exclude"
)

res_obj
```

## Collect Results

`cpmr` provides standardized accessors for downstream inspection.

```{r}
metrics <- collect_metrics(res_obj)
predictions <- collect_predictions(res_obj)
edges <- collect_edges(res_obj)

head(metrics)
head(predictions)
str(edges)
```

## Custom Resamples

You can pass a custom `resamples` list where each element stores assessment row
indices for one fold. The `resamples` list is validated to (a) have non-overlapping
assessment sets and (b) cover all complete-case rows exactly once. When
`na_action = "exclude"`, indices must refer to rows in the complete-case data.

```{r}
custom_resamples <- list(
  1:16,
  17:32,
  33:48,
  49:64,
  65:80
)

custom_res_obj <- fit_resamples(
  spec,
  conmat = conmat,
  behav = behav,
  resamples = custom_resamples,
  return_edges = "none"
)

collect_metrics(custom_res_obj)
```

## Next Steps

- See the leakage-focused vignette for covariate leakage control details.
- Use `summary()` and `tidy()` for reporting and post-processing.
