---
title: "Get Started with the Native CPM Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started with the Native CPM Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr`'s primary workflow is a native, matrix-first API built around
`cpm_spec()` together with `fit()` and `fit_resamples()`.

This vignette walks through:

- how to shape CPM inputs;
- how to fit a single native CPM model;
- how to read the helper-based `cpm_spec()` API;
- how to run leakage-safe resampling;
- how to inspect predictions and selected edges;
- how custom resamples and missing-data handling work.

## Prepare Inputs

CPM expects one row per subject and one column per edge.

- `conmat`: an `n x p` connectivity matrix;
- `behav`: a length-`n` numeric outcome vector;
- `covariates`: an optional `n x q` nuisance matrix.

```{r}
library(cpmr)

withr::local_seed(123)
n <- 80
p <- 200

conmat <- matrix(rnorm(n * p), nrow = n, ncol = p)
covariates <- matrix(rnorm(n), ncol = 1)
behav <- rowMeans(conmat[, 1:12, drop = FALSE]) +
  0.5 * covariates[, 1] +
  rnorm(n, sd = 0.5)
```

## Single Fit with `fit()`

Use `fit()` on a `cpm_spec()` object when you want a native single fit with
explicit CPM parameters.

```{r}
fit_obj <- fit(
  cpm_spec(
    selection = cpm_selection_cor(
      criterion = "p_value",
      level = 0.05
    )
  ),
  conmat = conmat,
  behav = behav,
  covariates = covariates
)

fit_obj
summary(fit_obj)
dim(fit_obj$edges)
```

For a single fit, `fit()` stores a `p x 2` edge-selection mask with `positive`
and `negative` columns by default. This is useful for inspecting the selected
network, but the performance reported by `summary()` is still in-sample.

## What `positive` and `negative` mean

The current API is easier to read once you attach a concrete meaning to those
two words:

- CPM first screens each edge against the outcome.
- `positive` edges are the screened edges whose association with the outcome is
  positive and passes the chosen threshold.
- `negative` edges are the screened edges whose association with the outcome is
  negative and passes the chosen threshold.
- For each subject, CPM then collapses each screened edge set into a
  subject-level summary feature, with optional edge weighting around the
  selection cutoff.

That gives a simple mental model:

- `positive_summary`: how strongly a subject expresses the screened
  outcome-positive edge set;
- `negative_summary`: how strongly a subject expresses the screened
  outcome-negative edge set.

In the current summary-based construction, `polarity` only decides how those
CPM-derived summaries are exposed to the second-stage outcome model:

- `polarity = "separate"` keeps both summaries. The `joint` stream fits a
  model with `positive_summary` and `negative_summary` together, while the
  `positive` and `negative` streams are single-summary diagnostics.
- `polarity = "net"` uses one
  `net_summary = positive_summary - negative_summary` feature and returns a
  single `net` stream.

## Build a Readable `cpm_spec()`

The current `cpm_spec()` API tries to keep the big modeling decisions visible
without forcing every setting into one long flat argument list.

A practical way to read a spec is:

- `selection = cpm_selection_cor(...)`: which correlation method is used to
  screen edges, how `level` is interpreted, and where the cutoff is set;
- `construction = cpm_construction_summary(...)`: how positive and negative
  screened edges become subject-level CPM predictors;
- `model = cpm_model_lm()`: how those CPM-derived predictors map to the final
  prediction.

Within `cpm_construction_summary(...)`:

- `polarity` decides whether positive and negative summaries stay separate or
  collapse to one net-summary feature;
- `weight_scale` decides whether edges contribute as a hard mask (`0`) or with
  sigmoid-style weights (`> 0`);
- `standardize_edges` decides whether each edge is fold-locally centered and
  scaled using the training data before the CPM summaries are computed.

This matches the edge-standardization step described by Rapuano et al. (2020):
edge strengths are z-scored within each training set and the corresponding
transformation is then applied to held-out data on that fold.

For example, this spec reads almost like a short sentence:

```{r}
readable_spec <- cpm_spec(
  selection = cpm_selection_cor(
    method = "spearman",
    criterion = "absolute",
    level = 0.1
  ),
  construction = cpm_construction_summary(
    polarity = "net",
    weight_scale = 0.03
  ),
  model = cpm_model_lm()
)

readable_spec
```

That means:

- rank-based screening with a direct absolute-correlation threshold;
- one `net_summary = positive_summary - negative_summary` feature;
- sigmoid-style edge weighting around the fold-local threshold;
- an explicit linear outcome model on the final network feature.

## Customize the CPM Strategy

`cpm_spec()` keeps the main CPM choices visible while using helpers for the
stages that naturally group settings together, such as selection and summary
construction.
In practice, most customization starts with just two questions:

- do you want a different edge-screening rule?
- do you want a different way to summarize or weight the network?

```{r}
rich_spec <- cpm_spec(
  selection = cpm_selection_cor(
    method = "spearman",
    criterion = "absolute",
    level = 0.1
  ),
  construction = cpm_construction_summary(
    polarity = "net",
    weight_scale = 0.03
  ),
  model = cpm_model_lm()
)

rich_fit <- fit(
  rich_spec,
  conmat = conmat,
  behav = behav,
  covariates = covariates
)

head(rich_fit$predictions)
```

Compared with the default spec, the example above changes four things:

- screening uses Spearman association instead of Pearson;
- screening uses a direct absolute-correlation cutoff instead of correlation
  p-value
  screening;
- the construction polarity is exposed as one `net` stream from
  `net_summary = positive_summary - negative_summary` instead of the
  classic `joint` / `positive` / `negative` trio;
- the second-stage model is made explicit, even though it still uses the
  default linear regression helper.

Prediction columns depend on `construction` polarity:

- `polarity = "separate"` yields `joint`, `positive`, and `negative`
  streams built from the positive and negative summaries;
- `polarity = "net"` yields a single `net` stream built from
  `net_summary = positive_summary - negative_summary`.

## Cross-Validated Resampling with `fit_resamples()`

Use `fit_resamples()` on the same `cpm_spec()` object for an out-of-sample
estimate of predictive performance.

```{r}
resample_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav,
  covariates = covariates,
  kfolds = 5
)

resample_obj
summary(resample_obj)
```

This keeps CPM-specific steps such as covariate handling, edge selection, and
model training inside each resample fold.

## Inspect Predictions and Edges

Native resampling results always keep raw observation-level predictions on the
object, while `summary()` gives an aggregate out-of-fold report with error
metrics as the default summary and correlations as supplementary output.

```{r}
predictions <- resample_obj$predictions

resample_summary <- summary(resample_obj)

resample_summary
head(resample_summary[["metrics"]])
head(predictions)
head(resample_metrics(resample_obj))
```

These three outputs have slightly different jobs:

- `summary(resample_obj)` prints the default aggregate report;
- `summary(resample_obj)[["metrics"]]` exposes the compact summary-level metric
  table stored on the summary object;
- `resample_metrics(resample_obj)` returns the raw pooled or fold-wise metric
  tables for further filtering, plotting, or export.

By default, `fit_resamples()` uses `return_edges = "none"` and skips edge
storage. This keeps the resampling object light when you only need predictive
performance.

If you also want fold-aggregated edge selection rates, request them explicitly:

```{r}
edge_resample_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav,
  covariates = covariates,
  kfolds = 5,
  return_edges = "sum"
)

dim(edge_resample_obj$edges)
```

`predictions` returns one row per original observation. If
`na_action = "exclude"` removed subjects before fitting, those rows are still
present and their `fold` value is `NA`.

When `return_edges = "sum"`, `cpmr` stores fold-summed edge counts for each
edge. If memory matters, keep the default `return_edges = "none"` or use
`return_edges = "all"` only when fold-wise edge arrays are truly needed.

As a rule of thumb:

- start with `return_edges = "none"` when you only care about prediction;
- use `return_edges = "sum"` when you want stable fold-aggregated edge
  selection rates;
- use `return_edges = "all"` only when you really need fold-level edge masks.

## Custom Resamples

If you already have a partition scheme, pass it through `resamples`.

```{r}
custom_resamples <- split(
  seq_len(n),
  cut(seq_len(n), breaks = 4, labels = FALSE)
)

custom_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav,
  resamples = custom_resamples,
  return_edges = "none"
)

summary(custom_obj)
```

Custom resamples must:

- be supplied as assessment-row indices;
- cover every complete-case subject exactly once;
- not overlap across folds.

When `na_action = "exclude"`, those indices still refer to original row numbers
of the complete-case subjects kept in the analysis.

## Missing-Data Handling

Set `na_action = "exclude"` if you want `cpmr` to fit on complete cases while
preserving original row positions in the outputs.

```{r}
behav_with_na <- behav
behav_with_na[c(3, 11)] <- NA_real_

na_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav_with_na,
  kfolds = 5,
  na_action = "exclude"
)

na_predictions <- na_obj$predictions
na_predictions[na_predictions$row %in% c(3, 11), ]
```

This behavior matters when you need to merge predictions back to subject-level
metadata after resampling.

## Next Steps

After you are comfortable with the native path:

- read the workflow-selection article for guidance on how to use the native
  `cpm_spec()` workflow in different settings;
- read the leakage-focused article if you need a more detailed covariate
  handling example;
- keep a `cpm_spec()` object around when you want an explicit reusable
  parameter object.
