---
title: "Native Getting Started with cpmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Native Getting Started with cpmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr`'s primary workflow is a native, matrix-first API built around
`spec()` together with `cpm()`.

This vignette walks through:

- how to shape CPM inputs;
- how to run leakage-safe resampling;
- how to inspect predictions and selected edges;
- how custom resamples and missing-data handling work.

## Prepare Inputs

CPM expects one row per subject and one column per edge.

- `conmat`: an `n x p` connectivity matrix;
- `behav`: a length-`n` numeric outcome vector;
- `covariates`: an optional `n x q` nuisance matrix.

```{r}
library(cpmr)

withr::local_seed(123)
n <- 80
p <- 200

conmat <- matrix(rnorm(n * p), nrow = n, ncol = p)
covariates <- matrix(rnorm(n), ncol = 1)
behav <- rowMeans(conmat[, 1:12, drop = FALSE]) +
  0.5 * covariates[, 1] +
  rnorm(n, sd = 0.5)
```

## Run CPM with `cpm()`

Use `cpm()` on a `spec()` object for an out-of-sample estimate of predictive
performance.

```{r}
resample_obj <- cpm(
  conmat = conmat,
  behav = behav,
  spec = spec(),
  covariates = covariates,
  resamples = 5
)

resample_obj
summary(resample_obj)
```

This keeps CPM-specific steps such as covariate handling, edge selection, and
model training inside each resample fold.

## Inspect Predictions and Edges

Native resampling results always keep raw observation-level predictions on the
object, while `summary()` gives an aggregate out-of-fold report with error
metrics as the default summary and correlations as supplementary output.

```{r}
predictions <- resample_obj$predictions

resample_summary <- summary(resample_obj)

resample_summary
head(resample_summary[["metrics"]])
head(predictions)
head(tidy(resample_obj, component = "metrics"))
head(tidy(resample_obj, component = "metrics", level = "pooled"))
```

These three outputs have slightly different jobs:

- `summary(resample_obj)` prints the default aggregate report;
- `summary(resample_obj)[["metrics"]]` exposes the compact summary-level metric
  table stored on the summary object;
- `tidy(resample_obj, component = "metrics")` returns the raw fold-wise metric
  table for further filtering, plotting, or export, while
  `level = "pooled"` returns pooled metrics.

By default, `cpm()` uses `return_edges = "sum"` and stores fold-aggregated edge
selection counts. This makes edge summaries immediately available in the common
resampling workflow.

If you want to control edge storage explicitly:

```{r}
edge_resample_obj <- cpm(
  conmat = conmat,
  behav = behav,
  spec = spec(),
  covariates = covariates,
  resamples = 5,
  return_edges = "sum"
)

dim(edge_resample_obj$edges)
```

`predictions` returns one row per original observation. If
`na_action = "exclude"` removed subjects before fitting, those rows are still
present and their `fold` value is `NA`.

When `return_edges = "sum"`, `cpmr` stores fold-summed edge counts for each
edge. If memory matters, switch to `return_edges = "none"` or use
`return_edges = "all"` only when fold-wise edge arrays are truly needed.

## Custom Resamples

If you already have a partition scheme, pass it through `resamples`.

```{r}
custom_resamples <- split(
  seq_len(n),
  cut(seq_len(n), breaks = 4, labels = FALSE)
)

custom_obj <- cpm(
  conmat = conmat,
  behav = behav,
  spec = spec(),
  resamples = custom_resamples,
  return_edges = "none"
)

summary(custom_obj)
```

Custom resamples must:

- be supplied as assessment-row indices;
- cover every complete-case subject exactly once;
- not overlap across folds.

When `na_action = "exclude"`, those indices still refer to original row numbers
of the complete-case subjects kept in the analysis.

## Missing-Data Handling

Set `na_action = "exclude"` if you want `cpmr` to fit on complete cases while
preserving original row positions in the outputs.

```{r}
behav_with_na <- behav
behav_with_na[c(3, 11)] <- NA_real_

na_obj <- cpm(
  conmat = conmat,
  behav = behav_with_na,
  spec = spec(),
  resamples = 5,
  na_action = "exclude"
)

na_predictions <- na_obj$predictions
na_predictions[na_predictions$row %in% c(3, 11), ]
```

This behavior matters when you need to merge predictions back to subject-level
metadata after resampling.

## Next Steps

After you are comfortable with the native path:

- read the workflow-selection article for guidance on how to use the native
  `spec()` + `cpm()` workflow in different settings;
- read the leakage-focused article if you need a more detailed covariate
  handling example;
- keep a `spec()` object around when you want an explicit reusable
  parameter object.
