---
title: "Native Getting Started with cpmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Native Getting Started with cpmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr`'s primary workflow is a native, matrix-first API built around
`cpm_spec()` together with `fit()` and `fit_resamples()`.

This vignette walks through:

- how to shape CPM inputs;
- how to fit a single native CPM model;
- how to run leakage-safe resampling;
- how to inspect predictions and selected edges;
- how custom resamples and missing-data handling work.

## Prepare Inputs

CPM expects one row per subject and one column per edge.

- `conmat`: an `n x p` connectivity matrix;
- `behav`: a length-`n` numeric outcome vector;
- `covariates`: an optional `n x q` nuisance matrix.

```{r}
library(cpmr)

withr::local_seed(123)
n <- 80
p <- 200

conmat <- matrix(rnorm(n * p), nrow = n, ncol = p)
covariates <- matrix(rnorm(n), ncol = 1)
behav <- rowMeans(conmat[, 1:12, drop = FALSE]) +
  0.5 * covariates[, 1] +
  rnorm(n, sd = 0.5)
```

## Single Fit with `fit()`

Use `fit()` on a `cpm_spec()` object when you want a native single fit with
explicit CPM parameters.

```{r}
fit_obj <- fit(
  cpm_spec(
    screen = cpm_screen(
      threshold = cpm_threshold("alpha", level = 0.05)
    )
  ),
  conmat = conmat,
  behav = behav,
  covariates = covariates
)

fit_obj
summary(fit_obj)
dim(fit_obj$edges)
```

For a single fit, `fit()` stores a `p x 2` edge-selection mask with `positive`
and `negative` columns by default. This is useful for inspecting the selected
network, but the performance reported by `summary()` is still in-sample.

## Customize the CPM Strategy

`cpm_spec()` keeps the main CPM choices visible while using helpers for the
parts that naturally come in pairs, such as screening and edge weighting.

```{r}
rich_spec <- cpm_spec(
  screen = cpm_screen(
    association = "spearman",
    threshold = cpm_threshold("effect_size", level = 0.1)
  ),
  network_summary = "difference",
  weighting = cpm_weighting("sigmoid", scale = 0.03),
  prediction_head = "linear_no_intercept"
)

rich_fit <- fit(
  rich_spec,
  conmat = conmat,
  behav = behav,
  covariates = covariates
)

head(rich_fit$predictions)
```

Prediction columns depend on `network_summary`:

- `network_summary = "separate"` yields `combined`, `positive`, and `negative`
  streams;
- `network_summary = "difference"` yields a single `difference` stream.

## Cross-Validated Resampling with `fit_resamples()`

Use `fit_resamples()` on the same `cpm_spec()` object for an out-of-sample
estimate of predictive performance.

```{r}
resample_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav,
  covariates = covariates,
  kfolds = 5
)

resample_obj
summary(resample_obj)
```

This keeps CPM-specific steps such as covariate handling, edge selection, and
model training inside each resample fold.

## Inspect Predictions and Edges

Native resampling results always keep raw observation-level predictions on the
object, while `summary()` gives an aggregate out-of-fold report with error
metrics as the default summary and correlations as supplementary output.

```{r}
predictions <- resample_obj$predictions

resample_summary <- summary(resample_obj)

resample_summary
head(resample_summary[["metrics"]])
head(predictions)
head(resample_metrics(resample_obj))
```

These three outputs have slightly different jobs:

- `summary(resample_obj)` prints the default aggregate report;
- `summary(resample_obj)[["metrics"]]` exposes the compact summary-level metric
  table stored on the summary object;
- `resample_metrics(resample_obj)` returns the raw pooled or fold-wise metric
  tables for further filtering, plotting, or export.

By default, `fit_resamples()` uses `return_edges = "none"` and skips edge
storage. This keeps the resampling object light when you only need predictive
performance.

If you also want fold-aggregated edge selection rates, request them explicitly:

```{r}
edge_resample_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav,
  covariates = covariates,
  kfolds = 5,
  return_edges = "sum"
)

dim(edge_resample_obj$edges)
```

`predictions` returns one row per original observation. If
`na_action = "exclude"` removed subjects before fitting, those rows are still
present and their `fold` value is `NA`.

When `return_edges = "sum"`, `cpmr` stores fold-summed edge counts for each
edge. If memory matters, keep the default `return_edges = "none"` or use
`return_edges = "all"` only when fold-wise edge arrays are truly needed.

## Custom Resamples

If you already have a partition scheme, pass it through `resamples`.

```{r}
custom_resamples <- split(
  seq_len(n),
  cut(seq_len(n), breaks = 4, labels = FALSE)
)

custom_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav,
  resamples = custom_resamples,
  return_edges = "none"
)

summary(custom_obj)
```

Custom resamples must:

- be supplied as assessment-row indices;
- cover every complete-case subject exactly once;
- not overlap across folds.

When `na_action = "exclude"`, those indices still refer to original row numbers
of the complete-case subjects kept in the analysis.

## Missing-Data Handling

Set `na_action = "exclude"` if you want `cpmr` to fit on complete cases while
preserving original row positions in the outputs.

```{r}
behav_with_na <- behav
behav_with_na[c(3, 11)] <- NA_real_

na_obj <- fit_resamples(
  cpm_spec(),
  conmat = conmat,
  behav = behav_with_na,
  kfolds = 5,
  na_action = "exclude"
)

na_predictions <- na_obj$predictions
na_predictions[na_predictions$row %in% c(3, 11), ]
```

This behavior matters when you need to merge predictions back to subject-level
metadata after resampling.

## Next Steps

After you are comfortable with the native path:

- read the workflow-selection article for guidance on how to use the native
  `cpm_spec()` workflow in different settings;
- read the leakage-focused article if you need a more detailed covariate
  handling example;
- keep a `cpm_spec()` object around when you want an explicit reusable
  parameter object.
