---
title: "CPM with Tidymodels Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CPM with Tidymodels Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr` exposes connectome-based predictive modeling (CPM) as a
tidymodels engine. This means we can:

- define CPM with `parsnip::cpm_reg()`;
- place it inside `workflows`;
- preprocess predictors with `recipes`;
- tune CPM thresholds with `tune`;
- evaluate performance with `yardstick`, including CPM-specific metrics.

The expected predictor shape is one row per subject and one column per
edge feature.

## Simulate A Small CPM Problem

```{r}
library(cpmr)
library(parsnip)
library(recipes)
library(rsample)
library(tune)
library(workflows)
library(yardstick)

set.seed(123)

n <- 90
p <- 50
x <- matrix(rnorm(n * p), nrow = n, ncol = p)
y <- x[, 1] - x[, 2] + 0.5 * x[, 3] + rnorm(n, sd = 0.5)

dat <- as.data.frame(x)
names(dat) <- paste0("edge_", seq_len(p))
dat$y <- y
```

## Fit A CPM Model

```{r}
spec <- cpm_reg(
  thresh_method = "sparsity",
  thresh_level = 0.2,
  network = "both"
)

fit_obj <- fit(spec, y ~ ., data = dat[1:70, ])
predict(fit_obj, dat[71:90, ])
```

The fitted engine stores the selected CPM edge masks. We can inspect
them with `collect_edges()`:

```{r}
collect_edges(fit_obj)
```

## Add Recipes And Workflows

Because `cpm_reg()` is a parsnip model specification, it composes
directly with workflows:

```{r}
rec <- recipe(y ~ ., data = dat[1:70, ]) |>
  step_normalize(all_predictors())

wf <- workflow() |>
  add_recipe(rec) |>
  add_model(spec)

wf_fit <- fit(wf, data = dat[1:70, ])
predict(wf_fit, dat[71:90, ])
```

## Tune Thresholds

CPM commonly varies by edge-selection threshold. The parsnip spec
supports that workflow directly:

```{r}
folds <- vfold_cv(dat[1:70, ], v = 3)

tuned_spec <- cpm_reg(
  thresh_method = "sparsity",
  thresh_level = tune(),
  network = "both"
)

tune_res <- workflow() |>
  add_formula(y ~ .) |>
  add_model(tuned_spec) |>
  tune_grid(
    resamples = folds,
    grid = tibble::tibble(thresh_level = c(0.15, 0.25, 0.35)),
    metrics = metric_set(rmse, cpm_cor, cpm_spearman)
  )

collect_metrics(tune_res)
```

## Collect Edges Across Resamples

`fit_resamples()` and `tune_grid()` do not retain fold-wise CPM edge masks
unless we explicitly extract them. The recommended workflow is:

1. use `extract_cpm_edges()` inside `control_resamples()` or
   `control_grid()`;
2. call `collect_edges()` on the resulting resample object.

```{r}
resample_ctrl <- control_resamples(
  save_pred = TRUE,
  extract = extract_cpm_edges
)

resample_res <- workflow() |>
  add_formula(y ~ .) |>
  add_model(spec) |>
  fit_resamples(
    resamples = folds,
    metrics = metric_set(rmse, cpm_cor),
    control = resample_ctrl
  )

collect_edges(resample_res, type = "sum")
collect_edges(resample_res, type = "prop")
```

For tuning results, `collect_edges()` keeps the tuning columns and `.config`
so we can compare edge stability across parameter settings:

```{r}
tune_ctrl <- control_grid(
  save_pred = TRUE,
  extract = extract_cpm_edges
)

tune_edges <- workflow() |>
  add_formula(y ~ .) |>
  add_model(tuned_spec) |>
  tune_grid(
    resamples = folds,
    grid = tibble::tibble(thresh_level = c(0.15, 0.25, 0.35)),
    metrics = metric_set(rmse, cpm_cor),
    control = tune_ctrl
  )

collect_edges(tune_edges, type = "sum")
```

## Grouped Resampling For Repeated Measures

If rows represent repeated scans or visits from the same subject, place the
subject identifier in the data frame, exclude it from the predictors, and use
`group_vfold_cv()` so the same subject never appears in both the analysis and
assessment portions of a split.

```{r}
subject_x <- x[rep(seq_len(30), each = 2), , drop = FALSE] +
  matrix(rnorm(60 * p, sd = 0.15), ncol = p)
grouped_dat <- data.frame(
  subject_id = factor(rep(paste0("subj_", seq_len(30)), each = 2)),
  y = rep(y[1:30], each = 2) + rnorm(60, sd = 0.2),
  subject_x,
  check.names = FALSE
)

group_folds <- group_vfold_cv(grouped_dat, group = subject_id, v = 3)

group_res <- workflow() |>
  add_formula(y ~ . - subject_id) |>
  add_model(spec) |>
  fit_resamples(
    resamples = group_folds,
    metrics = metric_set(rmse, cpm_cor),
    control = control_resamples(extract = extract_cpm_edges)
  )

collect_metrics(group_res)
collect_edges(group_res, type = "sum")
```

## Nested Cross-Validation Pattern

When CPM threshold tuning itself needs an outer unbiased assessment loop, use
`nested_cv()` and tune the workflow inside each outer split. The pattern below
is shown without evaluation to keep the vignette fast, but it matches the
workflow used in the test suite.

```{r, eval = FALSE}
nested_folds <- nested_cv(
  grouped_dat,
  outside = group_vfold_cv(group = subject_id, v = 3),
  inside = vfold_cv(v = 2)
)

outer_results <- lapply(seq_len(nrow(nested_folds)), function(i) {
  inner_res <- tune_grid(
    workflow() |>
      add_formula(y ~ . - subject_id) |>
      add_model(cpm_reg(
        thresh_method = "sparsity",
        thresh_level = tune(),
        network = "both"
      )),
    resamples = nested_folds$inner_resamples[[i]],
    grid = tibble::tibble(thresh_level = c(0.15, 0.25, 0.35)),
    metrics = metric_set(rmse, cpm_cor)
  )

  best <- select_best(inner_res, metric = "rmse")
  final_wf <- finalize_workflow(
    workflow() |>
      add_formula(y ~ . - subject_id) |>
      add_model(cpm_reg(
        thresh_method = "sparsity",
        thresh_level = tune(),
        network = "both"
      )),
    best
  )

  fit(final_wf, data = analysis(nested_folds$splits[[i]]))
})
```

## Leakage-Safe Resampling

CPM is a protocol rather than a single closed-form model. The package
keeps the CPM-specific steps inside each resampling split:

- edge selection is learned from training rows only;
- network aggregation is fit from training rows only;
- prediction for assessment rows uses only training-derived quantities.

This design keeps feature selection and performance evaluation separated
inside `fit_resamples()`-style tidymodels workflows.

## Where To Go Next

- Use `collect_edges()` to compare selected masks across different model
  fits.
- Add `recipes` preprocessing steps that do not leak assessment
  information.
- Combine `cpm_reg()` with nested resampling when selecting among many
  thresholds or workflow variants.
