---
title: "CPM with Tidymodels Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CPM with Tidymodels Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr` exposes connectome-based predictive modeling (CPM) as a
tidymodels engine. This means we can:

- define CPM with `parsnip::cpm_reg()`;
- place it inside `workflows`;
- preprocess predictors with `recipes`;
- tune CPM thresholds with `tune`;
- evaluate performance with `yardstick`, including CPM-specific metrics.

The expected predictor shape is one row per subject and one column per
edge feature.

## Simulate A Small CPM Problem

```{r}
library(cpmr)
library(parsnip)
library(recipes)
library(rsample)
library(tune)
library(workflows)
library(yardstick)

set.seed(123)

n <- 90
p <- 50
x <- matrix(rnorm(n * p), nrow = n, ncol = p)
y <- x[, 1] - x[, 2] + 0.5 * x[, 3] + rnorm(n, sd = 0.5)

dat <- as.data.frame(x)
names(dat) <- paste0("edge_", seq_len(p))
dat$y <- y
```

## Fit A CPM Model

```{r}
spec <- cpm_reg(
  thresh_method = "sparsity",
  thresh_level = 0.2,
  network = "both"
)

fit_obj <- fit(spec, y ~ ., data = dat[1:70, ])
predict(fit_obj, dat[71:90, ])
```

The fitted engine stores the selected CPM edge masks. We can inspect
them with `collect_edges()`:

```{r}
collect_edges(fit_obj)
```

## Add Recipes And Workflows

Because `cpm_reg()` is a parsnip model specification, it composes
directly with workflows:

```{r}
rec <- recipe(y ~ ., data = dat[1:70, ]) |>
  step_normalize(all_predictors())

wf <- workflow() |>
  add_recipe(rec) |>
  add_model(spec)

wf_fit <- fit(wf, data = dat[1:70, ])
predict(wf_fit, dat[71:90, ])
```

## Tune Thresholds

CPM commonly varies by edge-selection threshold. The parsnip spec
supports that workflow directly:

```{r}
folds <- vfold_cv(dat[1:70, ], v = 3)

tuned_spec <- cpm_reg(
  thresh_method = "sparsity",
  thresh_level = tune(),
  network = "both"
)

tune_res <- workflow() |>
  add_formula(y ~ .) |>
  add_model(tuned_spec) |>
  tune_grid(
    resamples = folds,
    grid = tibble::tibble(thresh_level = c(0.15, 0.25, 0.35)),
    metrics = metric_set(rmse, cpm_cor, cpm_spearman)
  )

collect_metrics(tune_res)
```

## Collect Edges Across Resamples

`fit_resamples()` and `tune_grid()` do not retain fold-wise CPM edge masks
unless we explicitly extract them. The recommended workflow is:

1. use `extract_cpm_edges()` inside `control_resamples()` or
   `control_grid()`;
2. call `collect_edges()` on the resulting resample object.

```{r}
resample_ctrl <- control_resamples(
  save_pred = TRUE,
  extract = extract_cpm_edges
)

resample_res <- workflow() |>
  add_formula(y ~ .) |>
  add_model(spec) |>
  fit_resamples(
    resamples = folds,
    metrics = metric_set(rmse, cpm_cor),
    control = resample_ctrl
  )

collect_edges(resample_res, type = "sum")
collect_edges(resample_res, type = "prop")
```

For tuning results, `collect_edges()` keeps the tuning columns and `.config`
so we can compare edge stability across parameter settings:

```{r}
tune_ctrl <- control_grid(
  save_pred = TRUE,
  extract = extract_cpm_edges
)

tune_edges <- workflow() |>
  add_formula(y ~ .) |>
  add_model(tuned_spec) |>
  tune_grid(
    resamples = folds,
    grid = tibble::tibble(thresh_level = c(0.15, 0.25, 0.35)),
    metrics = metric_set(rmse, cpm_cor),
    control = tune_ctrl
  )

collect_edges(tune_edges, type = "sum")
```

## Leakage-Safe Resampling

CPM is a protocol rather than a single closed-form model. The package
keeps the CPM-specific steps inside each resampling split:

- edge selection is learned from training rows only;
- network aggregation is fit from training rows only;
- prediction for assessment rows uses only training-derived quantities.

This design keeps feature selection and performance evaluation separated
inside `fit_resamples()`-style tidymodels workflows.

## Where To Go Next

- Use `collect_edges()` to compare selected masks across different model
  fits.
- Add `recipes` preprocessing steps that do not leak assessment
  information.
- Combine `cpm_reg()` with nested resampling when selecting among many
  thresholds or workflow variants.
