---
title: "Choosing a Workflow in cpmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing a Workflow in cpmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr` is organized around one native workflow: create a `spec()` object, then
run `cpm()` under the resampling plan that matches your question.

The main choice is not between different API families. It is between different
assessment plans:

- `cpm(..., resamples = NULL)` for leave-one-out CPM;
- `cpm(..., resamples = 5)` for k-fold CPM;
- `cpm(..., resamples = ...)` when you already have a partition scheme you want
  to enforce.

## Start with `spec()`

For most analyses, start with `spec()` and reuse it across CPM runs.

```{r}
library(cpmr)

set.seed(123)
conmat <- matrix(rnorm(80 * 200), nrow = 80)
behav <- rnorm(80)
model_spec <- spec()

loo_obj <- cpm(conmat = conmat, behav = behav, spec = model_spec)
resample_obj <- cpm(
  conmat = conmat,
  behav = behav,
  spec = model_spec,
  resamples = 5
)

summary(loo_obj)
summary(resample_obj)
head(summary(resample_obj)[["metrics"]])
```

This workflow is the best fit when you care about:

- large edge spaces;
- repeated resampling;
- leakage-safe covariate handling;
- future CPM-specific performance improvements such as fold-level reuse.

## Practical Rule of Thumb

Use `cpm()` when you want CPM results. Then choose the resampling plan that
matches your goal.

As a rule of thumb:

- use `summary(loo_obj)` when you want the default leave-one-out report;
- use `summary(resample_obj)` for the default aggregate k-fold report;
- use `summary(resample_obj)[["metrics"]]` for the compact summary metric
  table, or `tidy(resample_obj, component = "metrics")` when you need
  structured fold-wise or pooled metric tables instead of printed output.
