---
title: "Choosing a Workflow in cpmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing a Workflow in cpmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr` supports more than one way to express a CPM analysis, but they do not
all serve the same purpose.

The package is organized around a core-first design:

- the native `cpmr` path is the primary workflow for fitting, resampling, and
  future tuning extensions;
- `cpm_spec()` is the native way to make CPM parameters explicit and reusable.

## Recommended Path: `cpm_spec()`

For most analyses, start with `cpm_spec()` together with `fit()` and
`fit_resamples()`.

```{r}
library(cpmr)

set.seed(123)
conmat <- matrix(rnorm(80 * 200), nrow = 80)
behav <- rnorm(80)
spec <- cpm_spec()

fit_obj <- fit(spec, conmat = conmat, behav = behav)
resample_obj <- fit_resamples(spec, conmat = conmat, behav = behav, kfolds = 5)

summary(fit_obj)
summary(resample_obj)
```

This path is the best fit when you care about:

- large edge spaces;
- repeated resampling;
- leakage-safe covariate handling;
- future CPM-specific performance improvements such as fold-level reuse.

## Practical Rule of Thumb

Choose the native `cpm_spec()` path when you need speed, explicit CPM
behavior, or resampling-heavy experiments.
