---
title: "Choosing a Workflow in cpmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing a Workflow in cpmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr` is organized around one native workflow: create a `cpm_spec()` object,
then use `fit()` or `fit_resamples()` depending on the question you want to
answer.

The main choice is not between different API families. It is between:

- `fit()` for a single in-sample CPM fit;
- `fit_resamples()` for out-of-sample resampling;
- `fit_resamples(..., resamples = ...)` when you already have a partition
  scheme you want to enforce.

## Start with `cpm_spec()`

For most analyses, start with `cpm_spec()` and reuse it across single-fit and
resampling calls.

```{r}
library(cpmr)

set.seed(123)
conmat <- matrix(rnorm(80 * 200), nrow = 80)
behav <- rnorm(80)
spec <- cpm_spec()

fit_obj <- fit(spec, conmat = conmat, behav = behav)
resample_obj <- fit_resamples(spec, conmat = conmat, behav = behav, kfolds = 5)

summary(fit_obj)
summary(resample_obj)
head(summary(resample_obj)[["metrics"]])
```

This workflow is the best fit when you care about:

- large edge spaces;
- repeated resampling;
- leakage-safe covariate handling;
- future CPM-specific performance improvements such as fold-level reuse.

## Practical Rule of Thumb

Use `fit()` when you want to inspect one fitted CPM model on the current data.
Use `fit_resamples()` when you want an out-of-sample performance estimate or a
custom resampling design.

As a rule of thumb:

- use `summary(fit_obj)` or `tidy(fit_obj)` for single-fit summaries;
- use `summary(resample_obj)` for the default aggregate resampling report;
- use `summary(resample_obj)[["metrics"]]` or `resample_metrics(resample_obj)`
  when you need structured metric tables instead of printed output.
