---
title: "Choose and Reuse a CPM Specification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choose and Reuse a CPM Specification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`cpmr` is organized around one native workflow: create a `cpm_spec()` object,
then use `fit()` or `fit_resamples()` depending on the question you want to
answer.

The main choice is not between different API families. It is between:

- `fit()` for a single in-sample CPM fit;
- `fit_resamples()` for out-of-sample resampling;
- `fit_resamples(..., resamples = ...)` when you already have a partition
  scheme you want to enforce.

## Start with `cpm_spec()`

For most analyses, start with `cpm_spec()` and reuse it across single-fit and
resampling calls.

```{r}
library(cpmr)

set.seed(123)
conmat <- matrix(rnorm(80 * 200), nrow = 80)
behav <- rnorm(80)
spec <- cpm_spec()

fit_obj <- fit(spec, conmat = conmat, behav = behav)
resample_obj <- fit_resamples(spec, conmat = conmat, behav = behav, kfolds = 5)

summary(fit_obj)
summary(resample_obj)
head(summary(resample_obj)[["metrics"]])
```

If you are new to the helper-based spec, the easiest mental model is:

- `screen = cpm_screen(...)` decides how edges get into the model;
- `feature_space` decides how CPM-derived network strengths are exposed to the
  outcome model;
- `weighting = cpm_weighting(...)` decides whether edge contribution is binary
  or smooth;
- `model = cpm_model_lm()` and `standardize_edges` finish the modeling
  pipeline, with `standardize_edges` meaning fold-local edge standardization
  using the training data statistics. The default is `FALSE`, so the default
  spec stays closer to classic CPM unless you opt in explicitly.

In the current native workflow:

- screened positive-association edges define a `positive_strength`;
- screened negative-association edges define a `negative_strength`;
- `feature_space = "separate"` keeps those strengths available directly;
- `feature_space = "net"` uses one
  `net_strength = positive_strength - negative_strength` feature.

That means the classic `joint` stream is not a third strength. It is the
prediction stream that fits one model with both `positive_strength` and
`negative_strength` together.

That means you can usually start from `cpm_spec()` and only open up the part
you actually want to change, rather than rewriting the whole spec every time.

This workflow is the best fit when you care about:

- large edge spaces;
- repeated resampling;
- leakage-safe covariate handling;
- future CPM-specific performance improvements such as fold-level reuse.

## Practical Rule of Thumb

Use `fit()` when you want to inspect one fitted CPM model on the current data.
Use `fit_resamples()` when you want an out-of-sample performance estimate or a
custom resampling design.

As a rule of thumb:

- use `summary(fit_obj)` or `tidy(fit_obj)` for single-fit summaries;
- use `summary(resample_obj)` for the default aggregate resampling report;
- use `summary(resample_obj)[["metrics"]]` or `resample_metrics(resample_obj)`
  when you need structured metric tables instead of printed output.
